Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-09 eurocg/paper.tex
zwischenstand

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. In both cases we present polynomial-time dynamic programming algorithms. Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}. Let $ u \in V $ and $ \deg(u) = \delta $. 
Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment. We store the lengths of the possible stubs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs. We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.

@@

Here we assume that the intersection graph $ C = (V,E) $ is a rooted tree of maximum degree $ k $. We give a bottom up dynamic programming algorithm for solving \maxsped on $ C $. For a vertex $ u \in V $ let $ p(u) $ denote the parent of $ u $ in $ C $ and let $ c(u) $ denote the set of its children. For a vertex $ u \in V $ with parent $ p(u) \in V $ let $ a \in \mathbb{R}_+ $ be the shorter length between one of $ s(u) $'s endpoint and the intersection point between $ s(u) $ and $ s(p(u)) $. Now we can define the following two values, allowing us to categorize the stub-lengths into those not affecting the parents choice of stub and those which don't:

@@

\begin{align*}
	\solshort(u) = \max\{T_1(u),\dots,T_a(u)\} &&

@@

\end{align*}
It remains to define $ T_i(u) $ for a vertex $ u $ and stub length $ \ell_i(u) $:

@@

		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i \text{ intersects } s(v) \\

@@

By solving Recurrence~\ref{rec:tree} we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $.	Let $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $. Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $. Now let $ u \in V $ with degree $ \delta $ be an inner vertex. The values $ \solshort(v) $ and $ \sollong(v) $ are computed for all $ v \in c(u) $. $ T_i(u) $ for one $ 0 \leq i \leq \delta $ is then the sum of the maximum stubs we can draw among the children plus the stub length $ \ell_i(u) $ itself, which are just the values we are interested in. Setting then $ \sollong(u) $ and $ \solshort(u) $ as above gives the two necessary values for $ p(u) $.

@@

Solving Recurrence~\ref{rec:tree} straight forward would yield an algorithm running in $ O(mk^2) $. Using the order on the stub lengths though we can improve this to $ O(mk) $ by solving $ T_i(u) $ in linear time for one $ u \in V $.

@@

\begin{theorem}
	\label{thm:tree}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
%	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above. Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $. For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
%	
	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
\end{proof}

>>>>

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. In both cases we present polynomial-time dynamic programming algorithms. Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}. Let $ u \in V $ and $ \deg(u) = \delta $. 
Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment. We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs. We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.

@@

Here we assume that $ C =(V,E) $ is a rooted tree of maximum degree $ k $. 
We give a bottom up dynamic programming algorithm for solving \maxsped on $ C $. 
For each vertex $u \in V$ we compute and store the maximum ink values $T_i(u)$ for $i = 0, \dots, \delta$ with $\delta = \deg(u)$ for the subtree rooted at $u$ such that $s(u)$ is drawn as a pair of stubs of length $\ell_i(u)$.
For $ u \in V $ let $ p(u) $ denote the parent of $ u $ in $ C $ and let $ c(u) $ denote the set of its children. 
For $ u \in V $  %with parent $ p(u) \in V $ 
let $ i_p $ be the index of the stub length $\ell_{i_p}(u)$ induced by the intersection point of $s(u)$ and $s(p(u))$.
%the shorter length between one of $ s(u) $'s endpoints and the intersection point between $ s(u) $ and $ s(p(u)) $. 
We define the following two values, which allow us to categorize the stub lengths into those not affecting the stubs of the parent and those that do affect the parent:

@@

%\begin{align*}
\[	
	\solshort(u) = \max\{T_1(u),\dots,T_{i_p}(u)\} \qquad

@@

\]
%\end{align*}
Based on these two values we can recursively define $ T_i(u) $: % for a vertex $ u $ and stub length $ \ell_i(u) $:

@@

		\solshort(v) & \text{if } s(u) \text{ with length } \ell_i(u) \text{ intersects } s(v) \\

@@

The correctness of Recurrence~(\ref{rec:tree}) follows by induction. 
For a leaf $u$ in $C$ the set $c(u)$ is empty and the correctness of $T_i(u)$ is immediate.
Further, $\solshort(u) = T_1(u)$ and $\sollong(u) = T_0(u)$ are set correctly for the parent $p(u)$.
%Let $ C = (V,E) $ as above and $ u \in V $ a leaf of $ C $. Since $ c(u) $ is empty we simply get $ \solshort(u) = T_0(u) = \ell_1(u) $ and $ \sollong(u) = T_1(u) = \max\{\ell_0(u), \ell_1(u)\} $. 
For an inner vertex $ u $ with degree $\delta$ we can assume by the induction hypothesis that the values $ \solshort(v) $ and $ \sollong(v) $ are computed correctly for all children $ v \in c(u) $. 
Each value $ T_i(u) $ for $ 0 \leq i \leq \delta $ is then the stub length $ \ell_i(u) $ plus the sum of the maximum ink we can achieve among the children subject to the stubs of $u$ being drawn with length $\ell_i(u)$. 
Setting $ \sollong(u) $ and $ \solshort(u) $ as above yields the two maximum ink values that are relevant for $ p(u) $.

@@

Recurrence~(\ref{rec:tree}) can be solved naively in $ O(mk^2) $ time. 
Using the order on the stub lengths  we can improve this to $ O(mk) $ time by computing all $ T_i(u) $ for one $ u \in V $ in $O(k)$ time.
Let $u \in V$ be a vertex with degree $\deg(u) = \delta$. 
The values $T_0(u) = \ell_0(u) + \sum_{v\in c(u)} \solshort(v)$ and $T_1(u) = \ell_1(u) + \sum_{v\in c(u)} \sollong(v)$ for the whole segment $s(u)$ and the shortest stubs can be computed in $O(k)$ time each.

@@

\begin{theorem}
	\label{thm:tree}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
%	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above. Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $. For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
%

By solving Recurrence~(\ref{rec:tree}) we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.

	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
\end{proof}
